\emph{Cree}: A modern toolbox for readymade economic experiments

### 2015-06-23 Latex/paper.tex
Paper: Spelling and fomulation
Added abstract. Shorten som paragraphs. Combined small paragraphs into
larger coherent paragraphs. Added Table of Contents. Moved Installation
subsection to appendix.

\begin{abstract}
{\color{red} ...}
\end{abstract}
%\begin{keyword}Science \sep Publication \sep Complicated\end{keyword}

@@

@@

The great merit of \Cree is the very few restrictions it places on the equipment facing subjects. Few restrictions clear the way for much broader participation, strengthening the external validity of any experiment. \Cree experiments can be run in a myriad of settings, including laboratories, but also in \emph{extra-laboratory} settings such as classrooms, workplaces, or over the Internet \citep{Charness_Gneezy_Kuhn_2013}. With \Cree the computer-equipped laboratory is no longer a necessity -- lowering the overall costs of conducting economic experiments. The researcher can still choose to provide subjects with devices, but can just as easily let subjects use their own devices.

\Cree is build using web technologies. Content is structured using the markup language \emph{HTML}. The layout and presentation across different screen sizes is archived with the style sheet language \emph{CSS}. And all logic is constructed using the programming language \emph{JavaScript}. All web browsers interpret these three cornerstone languages and render pages accordingly. Subjects participate in \Cree experiments through a web browser. Because \Cree is fundamentally native to the web browser, it avoids many of the restrictions that other toolboxes suffer from.

In addition there is a vibrant ecosystem surrounding these web technologies. \Cree exploits this ecosystem and takes full advantage of the software libraries that exist. \ref{S:dependencies} lists all software libraries used by \Cree and gives a short description of each. Building on top of existing technologies provides stability, flexibility, and makes it easy to set up and run economic experiments. Fore instances, \Cree uses the software library \emph{Node.js} to run the server and handle networking issues. The asynchronous architecture of \emph{Node.js} gives \Cree the flexibility to handle real-time events, which many other toolboxes are not capable of. With this flexibility \Cree can run everything, from simple dictator game experiments, to highly sophisticated auction experiments. \Cree is a framework that provides the tools to handle many of the otherwise mundane, complicated or manually tasks required to set up and run economic experiments.

Technological advances opens up a new path, however the path is not without obstacles. This paper explores and discusses how to handle these obstacles. Some obstacles are alleviated through appropriate technical design of the toolbox. Other obstacles requires actions taken by the researcher.

While this paper deals with internet experiments in general, it continuously refers to studies that use subjects from Amazon's \emph{Mechanical Turk}. This is not to highlight or promote the use of \emph{Mechanical Turk}, it is simply because the bulk of internet experiments use this system.

@@

\emph{z-Tree} is currently the commonly used toolbox for conducting economic experiments. The development of \emph{z-Tree} started in 1995 and has been updated continuously \citep{Fischbacher_2007}. However the foundation and guiding principles of \emph{z-Tree} is based on the technology that was available two decades ago. Several other toolboxes exist that can run economic experiments. Below I divide the toolboxes into four categories based on the restrictions they place on the equipment facing subjects:

@@

Toolboxes, such as \emph{z-Tree}, are applications. They are written for a specific operating system, and will only run on that specific operating system. All toolboxes in this category require Microsoft Windows. These applications are not easily ported to other computer operating systems. And they will require a complete restructuring -- focused on optimising screen sizes and battery life -- if they are ever to run on smart-phones or tablets. In addition, applications will have to be downloaded (and installed) on the computer facing subjects, before they can participate in experiments. This requirement hinders subjects from using their personal devices.

@@

The fourth category of toolboxes run natively in the web browser, without the need for any browser plug-ins or predownloaded software. These toolboxes place the same low restrictions on the equipment facing subjects as \Cree. However, none of these toolboxes, except \emph{MWERT}, currently support simultaneous real-time interaction between subjects -- ie. it is impossible to design experiments where subjects continuously provide input, while they continuously and simultaneously receives feedback from the other subjects. Some toolboxes support sequential decision making, while others completely lack interaction among subjects. The toolbox \emph{MWERT} is build specifically to run real-time multiplayer experiments on the internet \citep{Hawkins_2014}. It relies on many of the same basic technologies as \Cree. However it has been built from scratch to a larger extent than \Cree, which instead relies heavily on existing software libraries. This fore instance imply that researchers will have to rewrite the core of the toolbox to run experiments where more than two subjects interact simultaneously. Whereas the number of subjects is simply a parameter to be specified in \Cree experiments. \emph{MWERT} has a single stage (not considering the stage with instructions and the exit-survey), which can then be repeated a specific number of types. Whereas researchers build a list with stages in \Cree. These stages can be identical but need not be, and in between each stage the researcher can add a synchronisation step or randomisation step, such that subjects wait for other subjects to catch up, or such that subjects are shuffled among groups, roles or treatments.

@@

The method used to recruit subjects will depend on the type of experiment and the desired representativeness of subjects.

It is commonly assumed that subjects are naïve, eg. encounter the experimental material for the first time, and have no a priori knowledge of the experiment, the purpose, nor knowledge of related experiments. This is not true for experiments conducted over the internet, and is often misapprehended by researchers, as shown by \citet*{Chandler_Mueller_Paolacci_2014}. They investigate experiments where subjects are recruited through Amazon's \emph{Mechanical Turk} system. \emph{Mechanical Turk} is a crowdsourcing platform of 500,000 workers that perform mostly small tasks and are paid a piece rate. Every worker has an account, and \emph{Mechanical Turk} prevents an account from performing the same task twice. \cite{Chandler_Mueller_Paolacci_2014} find that:

@@

\footnotetext[5]{\Cree does not test whether access to the server is blocked by a firewall, nor if the two required ports are open.}

@@

\footnotetext[6]{The JavaScript library \emph{jQuery} is used on 64.4\% of monitored websites. There is no statistics on the \emph{Bootstrap} CSS library, but its complementary JavaScript library is used on 7.7\% of websites. Source: \url{http://w3techs.com/technologies/overview/javascript_library/all}. While \emph{Node.js} is used as web-server on 0.1\% of monitored websites. Source: \url{http://w3techs.com/technologies/overview/web_server/all}.}

With numerous interconnected and simultaneously used devices, eventually a device will malfunction or the network will suffer a temporary disturbance. \Cree handles these types of errors by allowing subjects that briefly disconnects, to reconnect and continue the experiment. The researcher can specify how long connections are allowed to be absent, before a subject gets kick from the experiment -- the default is 10 seconds. Disconnects and reconnects are logged, so the researcher can later determine if it significantly impacted the results in the current stage of the experiment. I do not recommend reconnecting subjects after lengthy disconnects, since this would hold up the entire experiment and adversely impact the waiting time of the other participating subjects. In these instances the experiment proceeds without the disconnected subject. When one subject disconnects from a \Cree experiment, then the other subjects in the same group are redirected to a fallback page. This page informs the subjects about the fall out and explains why they are not able to continue the experiment being one man short. The remaining groups are unaffected by the disconnect, and \Cree is still able to  synchronise and randomise among the remaining subjects and groups. The researcher will have to manually calculate payments for subjects in the group that suffered a untimely disconnect, but \Cree is still able to calculate payments for subjects in the remaining groups.

@@

\footnotetext[7]{ie. when the subject is using a different tab in the web browser, when the web browser is minimised or fully obscured by another application, or when the device enters its power saving or screensaver mode (lock screen on smartphones). Source: \url{http://www.w3.org/TR/page-visibility/}}

@@

Laboratory experiment can often last an hour or more. There is very little literature on how unsupervised subjects perform in  experiments with long sessions.

Using \emph{instructional manipulation checks} \citet*{Oppenheimer_Meyvis_Davidenko_2009} show that student subjects pay less attention when unsupervised.

\cite{Goodman_Cryder_Cheema_2013} argue that \emph{Mechanical Turk} may not be appropriate for long or complicated experiments. They compare the results of their experiment -- which lasted about 16 minutes -- to that of \citet*{Paolacci_Chandler_Ipeirotis_2010} that only lasted around 5 minutes. Unlike \cite{Paolacci_Chandler_Ipeirotis_2010} they find that subjects from \emph{Mechanical Turk} are less attentive. They test attentiveness using \emph{instructional manipulation checks} at the end of the survey.

\citet*{Farrell_Grenier_Leiby_2014} run an hour-long experiment on \emph{Mechanical Turk} where subjects fulfil orders (sandwich orders or unspecified product orders). They find that \emph{Mechanical Turk} subjects exert the same amount of effort (or more), and make the same amount of mistakes (or fewer), than student participants. Even in the treatment where \emph{Mechanical Turk} subjects are paid one-fifth of the student participants, where tasks are less intrinsically interesting, and where their wage rate is independent of performance (flat rate of \$5).

The results of \cite{Farrell_Grenier_Leiby_2014} may be due to the selection process, ie. the \emph{Mechanical Turk} subjects that volunteer for their experiment, are more motivated by financial incentives, than other individuals. An indication of this is that 73\% report `extra income' and 28\% report `entertainment' as reasons for using \emph{Mechanical Turk}. Whereas in the \cite{Paolacci_Chandler_Ipeirotis_2010} study it is respectively 61\% and 41\%. Similarly in a short survey by \cite{Williamson_2014} on \emph{Mechanical Turk} only 28.9\% respond that they would be willing to participate in a hour-long follow-up interview for an additional \$15.

@@

There are two alternative approaches, regarding time and duration that the researchers might consider when designing the experiment. One is the approach of \citet*{Pettit_Friedman_Kephart_Oprea_2014} which speeds up the adjustment process by allowing subjects to continuously change and adapt strategies. They refer to one study where distinctive, long-term implications are observable in rounds lasting only 120 seconds. The other approach, suggested by \cite{Charness_Gneezy_Kuhn_2013}, is to create long time-series data, ie. by recalling subjects every day, week or month. In a traditional laboratory experiment, this would be time consuming and highly inconvenient for subjects, but less so in extra-laboratory settings such as classrooms or over the internet. \cite[p. 96]{Charness_Gneezy_Kuhn_2013} argue that \emph{``the classic 10–20 period repetition of the public-goods game may not be a good analog for the expression of preference for tax policy that citizens make when they vote every couple of years''}. \Cree can execute both alternative approaches, since it allows real-time interaction among subjects and works in \emph{extra-laboratory} settings.

@@

For in-person experiments the researcher can pay subjects in cash. For experiments conducted over the internet payment will require a third-party payment provider. This could be national provides, such as the Danish \emph{MobilePay} or \emph{Swipp}. Or it could be international providers such as \emph{PayPal}. When a third-party payment provider is used, subjects will enter their mobile phone number, email address, or receive a voucher code they can use to claim their payment. At the end of every session \Cree generates a list with the payment for every subject, and with the respective payment method and transfer details.

@@

Subjects may have questions regarding instructions or procedures during the experiment. In supervised in-person experiments, the experimenter can walk around and answer these questions. For experiments conducted over the internet, \Cree offers the researcher the option to include a real-time chat function at specific stages of the experiment. Using this chat function subjects can privately ask the experimenter questions and receive answers. However this should be a fallback option reserved for special cases, as it may lead to unintended \emph{experimenter effects}, ie. subjects trying to please the experimenter, rather than behaving intrinsically. Instead researchers should first take all the necessary provisions \citep{Guala_2005}; insure that instructions, and procedures are clear and easily understandable, run trail experiments, and if necessary adjust instructions or simplify the experimental design. Alternatively the researcher can use instructional checks, ie. test if the subjects correctly understood instructions, before allowing them to proceed further in the experiment.

@@

Discouraging cheating in unsupervised experiments is difficult. And additional considerations are needed since subjects participate in experiments using a web browser. First, \cite{Goodman_Cryder_Cheema_2013} have shown that \emph{Mechanical Turk} subjects search for the correct answers to factual questions on the internet. Subjects look up answers, even when payment is independent of answering correctly. And when subjects are paid for the correct answer, they are more likely to look up the answer. Asking subjects not to use any externals sources reduces the number of subjects that look up the answer, but it does not completely eliminate cheating. Secondly, toolboxes that are native to the web browser are at a disadvantage compared to the toolboxes in the three other categories (discussed in section~\ref{SS:Alternatives}), since pages are rendered in the subject's web browser. Thus the source code to the unrendered pages is easily viewable and relatively easy to manipulate \citep[25m. 12s.]{Hawkes_2011}. In \Cree Subjects cannot cheat by manipulating how the page is rendered, since the server has final authority and controls the experiment. But subjects may attempt to cheat by sending false or manipulated inputs back to the server. It is therefore crucial, when the researcher designs the experiment, that all inputs are validated on the server. In addition since the source code of the page is accessible, it should not contain information that disclose the purpose of the experiment, the treatment, nor optimal strategies -- eg. instead of using variables names such as \texttt{Option\_Nash\_Equilibrium} and \texttt{Option\_Sub\_Optimal}, the researcher should use generic names such as \texttt{OptionX} and \texttt{OptionY}. A third issue is coercion among subjects. Coercion may arise in unsupervised in-person experiments where subjects are seated closely (ie. classrooms, work places, etc.), or in internet experiments where subjects are able to self-select into the experiment, and therefore invite others to join with the aim of manipulating the experiment. In \Cree the researcher can randomise subjects across groups, roles and treatments, but other than this \Cree cannot alleviate coercion. If possible it is recommended that in-person experiments be supervised. And researchers should limit the subject's ability to self-select into the experiment, when conducting experiments over the internet.

@@

Cree is installed on a computer with Node.js using its package manager (\emph{npm}). To install Cree open \emph{Command Prompt} on Windows, or \emph{Terminal} on Mac/Linux, enter \texttt{npm~install~Cree} and press enter.

@@

The guiding principle behind \Cree is to offer a solid and highly flexible foundation that gives researchers the necessary tools to carry out their research, irrespective of location and available devices. The current version of \Cree does require some programming skills. An interface -- providing convenience when running preprogrammed experiments and when implementing minor changes -- could be added on top of this foundation in a future version. The interface would be a supplement. But with the foundation in order, this would not be at the expense of the general flexibility of \Cree. \Cree relies on JavaScript to build logic into experiments. Because JavaScript is a general-purpose programming language and widely used, many of the procedures necessary to build more complex experiments are already easily available. Thereby avoiding the trap noted by \citet[p. 177]{Fischbacher_2007} when discussing the current limitations of \emph{z-tree}; \emph{``there are situations in which programming is unnecessarily complicated, since z-Tree neither knows procedures nor complex data types ... One of the planned future steps in the development of z-Tree is thus an extension of the programming language to make programming of more complex experiments easier''}. \Cree uses the CSS framework \emph{Bootstrap}, which insures that pages scale seamlessly across small and big screens. In addition it ensures that pages look identical across different browsers and devices, for given screen sizes. Since Bootstrap handles all this, knowledge of CSS is not need to build elegant experiments that have a consistent look across devices. Some knowledge of HTML is  helpful, but not required since the researcher can use third-party drag and drop tools such as \emph{LayoutIt!} or \emph{Jetstrap} to build pages that work with \emph{Bootstrap}\footnotemark[8].

@@

This paper introduces \Cree, a modern toolbox to design and conduct economic experiments. \Cree places very few restrictions on the devices facing subjects, thus relinquishing economic experiments from the grips of the computer-equipped laboratory. In turn economic experiments designed in \Cree can also run in \emph{extra-laboratory} settings such as classrooms, workplaces and over the internet. While \Cree provides researchers with new tools and opportunities, it also opens up for new issues and obstacles. This paper shows how many of these obstacles are solved in the technical design of \Cree, while a few obstacles remain that require action by the researcher. The main remaining issue facing researchers when transferring from a laboratory setting to an \emph{extra-laboratory} setting is not subject attention, subject compensation, nor the quality of the results. It is recruitment and specifically the selection process. If the experiment is designed appropriately the first set of issues resolve themselves. However the method used to recruit subjects, especially if subjects self-select into the experiment, may change the mindset, behaviour, validity and ultimately the results of the experiment. The selection process requires a third-party prescreening system to manage the pool of subjects. And it may require running prescreen surveys to appropriately exclude subjects on a priori grounds. The problems surrounding the selection process are not exclusive to internet experiments, just more prevalent in this setting.

@@

There is a vibrant ecosystem and many general-purpose software libraries that surround the three web technologies HTML, CSS and JavaScript. \Cree exploits this ecosystem and takes full advantage of the software libraries that exist. This provides stability, flexibility, and makes it easy to set up and run economic experiments. This section briefly describes each of the software libraries, which \Cree depends upon.

>>>>

\begin{abstract}
{\Cree is a toolbox to design and run economic experiments. \Cree places very few restrictions on the devices facing subjects, thus \Cree experiments can be run in a myriad of settings. This opens up a new path for researchers, however the path is not without obstacles. This paper describes how many of these obstacles are alleviated though the technical design of \Cree. The main remaining obstacle is the selection process, especially if subjects can self-select into experiments.}
\end{abstract}
%\begin{keyword}Science \sep Publication \sep Complicated\end{keyword}

@@

% \tableofcontents
\singlespacing
\setcounter{tocdepth}{1}
\makeatletter
	\@starttoc{toc}
\makeatother
\onehalfspacing

@@

The great merit of \Cree is the very few restrictions it places on the equipment facing subjects. Few restrictions clear the way for much broader participation, strengthening the external validity of any experiment. \Cree experiments can run in a myriad of settings, including laboratories, but also in \emph{extra-laboratory} settings such as classrooms, workplaces, or over the Internet \citep{Charness_Gneezy_Kuhn_2013}. With \Cree the computer-equipped laboratory is no longer a necessity -- lowering the overall costs of conducting economic experiments. The researcher can still choose to provide subjects with devices, but can just as easily let subjects use their own devices.

\Cree is build using web technologies. Content is structured using the markup language \emph{HTML}. The layout and presentation across different screen sizes is archived with the style sheet language \emph{CSS}. And all logic is constructed using the programming language \emph{JavaScript}. All web browsers interpret these cornerstone languages and render pages accordingly. Subjects participate in \Cree experiments through a web browser. Because \Cree is fundamentally native to the web browser, it avoids many of the restrictions that other toolboxes suffer from.

In addition there is a vibrant ecosystem surrounding these web technologies. \Cree exploits this ecosystem and takes full advantage of the software libraries that exist. \ref{S:dependencies} lists all the libraries used by \Cree and gives a short description of each. Building on top of existing technologies provides stability, flexibility, and makes it easy to set up and run economic experiments in \Cree. Fore instances, \Cree uses the software library \emph{Node.js} to run the server and handle networking issues. The asynchronous architecture of \emph{Node.js} gives \Cree the flexibility to handle real-time events, which many other toolboxes are not capable of. With this flexibility \Cree can run everything, from simple dictator game experiments, to highly sophisticated auction experiments. \Cree is a framework that provides the tools to handle many of the otherwise mundane, complicated or manually tasks required to set up and run economic experiments.

Technological advances opens up a new path, however the path is not without obstacles. This paper explores and discusses how to handle these obstacles. Some obstacles are alleviated through appropriate technical design of the toolbox. Other obstacles requires actions taken by the researcher. While this paper deals with internet experiments in general, it continuously refers to studies that use subjects from Amazon's \emph{Mechanical Turk}. This is not to highlight or promote the use of \emph{Mechanical Turk}, it is simply because the bulk of internet experiments use this system.

@@

\emph{z-Tree} is currently the commonly used toolbox for conducting economic experiments. The development of \emph{z-Tree} started in 1995 and has been updated continuously \citep{Fischbacher_2007}. However the foundation and guiding principles of \emph{z-Tree} is based on the technology that was available two decades ago. Several other toolboxes exist that run economic experiments. Below I divided the toolboxes into four categories based on the restrictions they place on the equipment facing subjects:

@@

Toolboxes, such as \emph{z-Tree}, are applications. They are written for a specific operating system, and will only run on that specific operating system. All toolboxes in this category require Microsoft Windows. These applications are not easily ported to other computer operating systems. And they will require a complete restructuring -- focused on optimising screen sizes and battery life -- if they are ever to run on smart-phones or tablets. In addition, these applications need to be downloaded (and installed) on the computer, before subjects can participate. This requirement hinders subjects from using their personal devices.

@@

The fourth category of toolboxes run natively in the web browser, without the need for any browser plug-ins or predownloaded software. These toolboxes place the same low restrictions on the equipment facing subjects as \Cree. However, none of these toolboxes, except \emph{MWERT}, currently support simultaneous real-time interaction between subjects -- ie. it is impossible to design experiments where subjects continuously provide input, while they continuously and simultaneously receives feedback from the other subjects. Some toolboxes support sequential decision making, while others completely lack interaction among subjects. The toolbox \emph{MWERT} is build specifically to run real-time multiplayer experiments on the internet \citep{Hawkins_2014}. It relies on many of the same basic technologies as \Cree, however \emph{MWERT} takes less advantages of the availability of general-purpose software libraries, making it more cumbersome to develop experiments. Fore instance the ``build for scratch'' approach manifests itself when the researcher wants to run experiments where more than two subjects interact simultaneously. In \emph{MWERT} the researcher has to rewrite the core of the toolbox, whereas the number of subjects is simply a parameter to be specified in \Cree experiments. Another limitation is the structure of the stages. \emph{MWERT} has just one stage (disregarding the instructions and the exit-survey). Subjects repeat this single stage a given number of times, whereafter the session ends. \Cree on the other hand gives researchers the flexibility to build a list of all the stages. These stages can be identical, but need not be. And in between each stage, the researcher can add a synchronisation step such that subjects wait for other subjects to catch up, or they can add a randomisation step such that subjects are shuffled among groups, roles or treatments.

@@

The method used to recruit subjects will depend on the type of experiment and the desired representativeness of subjects. It is commonly assumed that subjects are naïve, eg. encounter the experimental material for the first time, and have no a priori knowledge of the experiment, the purpose, nor knowledge of related experiments. This is not true for experiments conducted over the internet, and is often misapprehended by researchers, as shown by \citet*{Chandler_Mueller_Paolacci_2014}. They investigate experiments where subjects are recruited through Amazon's \emph{Mechanical Turk} system. \emph{Mechanical Turk} is a crowdsourcing platform of 500,000 workers that perform mostly small tasks and are paid a piece rate. Every worker has an account, and \emph{Mechanical Turk} prevents an account from performing the same task twice. \cite{Chandler_Mueller_Paolacci_2014} find that:

@@

\footnotetext[5]{\Cree does not test if access to the server is blocked by a firewall, nor if the two required ports are open.}

@@

\footnotetext[6]{\emph{jQuery} is used on 64\% of monitored websites. There is no statistics on the \emph{Bootstrap} CSS library, but its complementary JavaScript library is used on 8\% of websites. Source: \url{http://w3techs.com/technologies/overview/javascript_library/all}. While \emph{Node.js} is used as web-server on 0.1\% of monitored websites. Source: \url{http://w3techs.com/technologies/overview/web_server/all}.}

With numerous interconnected and simultaneously used devices, eventually a device will malfunction or the network will suffer a temporary disturbance. \Cree handles these types of errors by allowing subjects that briefly disconnects, to reconnect and continue the experiment. The researcher can specify how long connections are allowed to be absent, before a subject gets kick from the experiment -- the default is 10 seconds. Disconnects and reconnects are logged, so the researcher can later determine if it significantly impacted the results of the experiment. I do not recommend reconnecting subjects after lengthy disconnects, since this would hold up the entire experiment and adversely impact the waiting time of the other participating subjects. In these instances the experiment proceeds without the disconnected subject. When a subject definitively disconnects from a \Cree experiment, then the other subjects in the same group are redirected to a fallback page. This page informs the subjects about the fall out and explains why they are not able to continue the experiment being one man short. The remaining groups are unaffected by the disconnect, and \Cree is still able to  synchronise and randomise among the remaining subjects and groups. The researcher will have to manually calculate payments for subjects in the group that suffered a untimely disconnect, but \Cree is still able to calculate payments for subjects in the remaining groups.

@@

\footnotetext[7]{ie. when the subject uses a different tab in the web browser, when the browser is minimised or fully obscured by another application, or when the device enters its power saving mode (lock screen on smartphones). Source: \url{http://www.w3.org/TR/page-visibility/}}

@@

Laboratory experiment can often last an hour or more. There is very little literature on how unsupervised subjects perform in  experiments with long sessions. Using \emph{instructional manipulation checks} \citet*{Oppenheimer_Meyvis_Davidenko_2009} show that student subjects pay less attention when unsupervised. \cite{Goodman_Cryder_Cheema_2013} argue that \emph{Mechanical Turk} may not be appropriate for long or complicated experiments. They compare the results of their experiment -- which lasted about 16 minutes -- to that of \citet*{Paolacci_Chandler_Ipeirotis_2010} that only lasted around 5 minutes. Unlike \cite{Paolacci_Chandler_Ipeirotis_2010} they find that subjects from \emph{Mechanical Turk} are less attentive. They test attentiveness using \emph{instructional manipulation checks} at the end of the survey. \citet*{Farrell_Grenier_Leiby_2014} actually run an hour-long experiment on \emph{Mechanical Turk} where subjects fulfil orders (sandwich orders or unspecified product orders). They find that \emph{Mechanical Turk} subjects exert the same amount of effort (or more), and make the same amount of mistakes (or fewer), than student participants. Even in the treatment where \emph{Mechanical Turk} subjects are paid one-fifth of the student participants, where tasks are less intrinsically interesting, and where their wage rate is independent of performance (flat rate of \$5). The results of \cite{Farrell_Grenier_Leiby_2014} may be due to the selection process, ie. the \emph{Mechanical Turk} subjects that volunteer for their experiment, are more motivated by financial incentives, than other individuals. An indication of this is that 73\% report `extra income' and 28\% report `entertainment' as reasons for using \emph{Mechanical Turk}. Whereas in the \cite{Paolacci_Chandler_Ipeirotis_2010} study it is respectively 61\% and 41\%. Similarly in a short survey by \cite{Williamson_2014} on \emph{Mechanical Turk} only 28.9\% respond that they would be willing to participate in a hour-long follow-up interview for an additional \$15.

@@

There are two alternative approaches, regarding time and duration that the researchers might consider when designing the experiment. One is the approach of \citet*{Pettit_Friedman_Kephart_Oprea_2014} which speeds up the adjustment process by allowing subjects to continuously change and adapt strategies. They refer to one study where distinctive, long-term implications are observable in rounds lasting only 120 seconds. The other approach, suggested by \cite{Charness_Gneezy_Kuhn_2013}, is to create long time-series data, ie. by recalling subjects every day, week or month. In a traditional laboratory experiment, this would be time consuming and highly inconvenient for subjects, but less so in extra-laboratory settings such as classrooms or over the internet. \citet[p. 96]{Charness_Gneezy_Kuhn_2013} argue that \emph{``the classic 10–20 period repetition of the public-goods game may not be a good analog for the expression of preference for tax policy that citizens make when they vote every couple of years''}. \Cree can execute both alternatives, since it allows real-time interaction among subjects and works in \emph{extra-laboratory} settings.

@@

For in-person experiments the researcher can pay subjects in cash. For experiments conducted over the internet payment will require a third-party payment provider. This could be national provides, such as the Danish \emph{MobilePay} or \emph{Swipp}. Or it could be international providers such as \emph{PayPal}. When a third-party payment provider is used, subjects will enter their mobile phone number, email address, or receive a voucher code they can use to claim their payment. At the end of every session \Cree generates a list with each subject's payment, preferred method of payment, and transfer details.

@@

Subjects may have questions regarding instructions or procedures during the experiment. In supervised in-person experiments, the experimenter can walk around and answer these questions. For experiments conducted over the internet, \Cree can be extend to include a real-time chat function. Using this chat function subjects can privately ask the experimenter questions and receive answers. However this should be a fallback option reserved for special cases, as it may lead to unintended \emph{experimenter effects}, ie. subjects trying to please the experimenter, rather than behaving intrinsically. %Instead researchers should first take all the necessary provisions \citep{Guala_2005}; insure that instructions, and procedures are clear and easily understandable, run trail experiments, and if necessary adjust instructions or simplify the experimental design. Alternatively the researcher can use instructional checks, ie. test if the subjects correctly understood instructions, before allowing them to proceed further in the experiment.

@@

Discouraging cheating in unsupervised experiments is difficult. And additional case is needed since subjects participate in experiments using a web browser. First, \cite{Goodman_Cryder_Cheema_2013} have shown that \emph{Mechanical Turk} subjects search for the correct answers to factual questions on the internet. Subjects look up answers, even when payment is independent of answering correctly. And when subjects are paid for the correct answer, they are more likely to look up the answer. Asking subjects not to use any externals sources reduces the number of subjects that look up the answer, but it does not completely eliminate cheating. Secondly, toolboxes that are native to the web browser are at a disadvantage compared to the toolboxes in the three other categories (discussed in section~\ref{SS:Alternatives}), since pages are rendered in the subject's web browser. Thus the source code to the unrendered pages is easily viewable and relatively easy to manipulate \citep[25m. 12s.]{Hawkes_2011}. In \Cree Subjects cannot cheat by manipulating how the page is rendered, since the server has final authority and controls the experiment. But subjects may attempt to cheat by sending false or manipulated inputs back to the server. It is therefore crucial, when the researcher designs the experiment, that all inputs are validated on the server. In addition since the source code of the page is accessible, it should not contain information that disclose the purpose of the experiment, the treatment, nor optimal strategies -- eg. instead of using variables names such as \texttt{Option\_Nash\_Equilibrium} and \texttt{Option\_Sub\_Optimal}, the researcher should use generic names such as \texttt{OptionX} and \texttt{OptionY}. A third issue is coercion among subjects. Coercion may arise in unsupervised in-person experiments where subjects are seated closely (ie. classrooms, work places, etc.), or in internet experiments where subjects are able to self-select into the experiment, and therefore invite others to join with the aim of manipulating the experiment. In \Cree the researcher can randomise subjects across groups, roles and treatments, but other than this \Cree cannot alleviate coercion. If possible it is recommended that in-person experiments be supervised. And researchers should limit the subject's ability to self-select into the experiment, when conducting experiments over the internet.

@@

@@

The guiding principle behind \Cree is to offer a solid and highly flexible foundation that gives researchers the necessary tools to carry out their research, irrespective of location and available devices. The current version of \Cree does require some programming skills. An interface -- providing convenience when running preprogrammed experiments and when implementing minor changes -- could be added on top of this foundation in a future version. The interface would be a supplement. But with the foundation in order, this would not be at the expense of the general flexibility of \Cree. \Cree relies on JavaScript to build logic into experiments. Because JavaScript is a general-purpose programming language and widely used, many of the procedures necessary to build more complex experiments are already easily available. Thereby avoiding the trap noted by \citet[p. 177]{Fischbacher_2007} when discussing the current limitations of \emph{z-tree}; \emph{``there are situations in which programming is unnecessarily complicated, since z-Tree neither knows procedures nor complex data types ... One of the planned future steps in the development of z-Tree is thus an extension of the programming language to make programming of more complex experiments easier''}. \Cree uses the CSS framework \emph{Bootstrap}, which insures that pages scale seamlessly across small and big screens. In addition it ensures that pages look identical across different browsers and devices, for given screen sizes. Since Bootstrap handles all this, knowledge of CSS is not need to build elegant experiments that have a consistent look across devices. Some knowledge of HTML is  helpful, but not required since the researcher can use third-party drag and drop tools such as \emph{LayoutIt!} or \emph{Jetstrap} to build pages that work with \emph{Bootstrap}\footnotemark[8]. \ref{S:dictator} gives a broad outline how a Dictator game experiment would look in \Cree.

@@

This paper introduces \Cree, a modern toolbox to design and conduct economic experiments. \Cree places very few restrictions on the devices facing subjects, thus relinquishing economic experiments from the grips of the computer-equipped laboratory. In turn economic experiments designed in \Cree can also run in \emph{extra-laboratory} settings such as classrooms, workplaces and over the internet. While \Cree provides researchers with new tools it also opens up for new issues and obstacles. This paper shows how many of these obstacles are solved in the technical design of \Cree, while a few obstacles remain that require action by the researcher. The main remaining issue facing researchers when transferring from a laboratory setting to an \emph{extra-laboratory} setting is not subject attention, subject compensation, nor the quality of the results. It is recruitment and specifically the selection process. If subjects self-select into the experiment, it may change the mindset, behaviour, validity and ultimately the results. A third-party prescreening system is required to manage the pool of subjects. And running prescreen surveys to appropriately exclude subjects on a priori grounds might also. % be required. The problems surrounding the selection process are not exclusive to internet experiments, just more prevalent in this setting.

@@

There is a vibrant ecosystem and many general-purpose software libraries surrounding the three web technologies; HTML, CSS and JavaScript. \Cree exploits this ecosystem and takes advantage of these libraries. This increases the stability of \Cree, provides flexibility, and makes it easy to conduct economic experiments. This section briefly describes each of the software libraries, which \Cree depends upon.

\begin{description}
  \item[Node.js:] \Cree uses the software library \emph{Node.js} to run the server and control the experiment. \emph{Node.js} is build on the same JavaScript engine as the Chrome browser, but \emph{Node.js} runs JavaScript on the server. This means that \Cree uses the programming language JavaScript to induce logic, both at the server level, and at the web browser level. \emph{Node.js} has a asynchronous architecture -- so instead of executing the code from the top, one line after the other, \emph{Node.js} binds all code to events and the code is only executed once an event happen (eg. when subjects enter the experiment, or when subjects sends messages, etc.). This asynchronous architecture gives \Cree the flexibility to effortlessly handle real-time events. \emph{Node.js} is easily installed from \url{http://nodejs.org}, and comes with the package manager \emph{npm}.
  \item[cloak:] This \emph{Node.js} module or package aims to provide a network layer to HTML games. \Cree uses \emph{cloak} to handle all networking issues, e.g. connecting subjects, queuing subjects, sending messages back and forth between subjects, etc. See \url{http://incompl.github.io/cloak/}
  \item[knuth-shuffle:] \Cree uses this \emph{Node.js} module to randomise subjects across groups, roles and treatments respectively. The underlying algorithm is known as the \emph{Fisher–Yates shuffle} or the \emph{Knuth shuffle}. And the algorithm is unbiased. See \url{https://github.com/coolaj86/knuth-shuffle}
  \item[connect:] This is the \emph{Node.js} module that \Cree uses to serve web pages to subjects. See \url{https://github.com/senchalabs/connect}
  \item[ua-parser-js:] \Cree uses this \emph{Node.js} module to identify and block specific devices and web browsers. See \url{http://faisalman.github.io/ua-parser-js/}
  \item[underscore:] This is a \emph{Node.js} module that extends JavaScript and provides additional useful functions. See \url{http://underscorejs.org}
  \item[ipify.org:] \Cree calls this \emph{application programming interface} (API), and it returns the public URL address, ie. the URL address that subjects use when joining the experiment over the internet. See \url{https://www.ipify.org}.
  \item[forever:] This command line tool gives \Cree its fully-continuous option described in section~\ref{SS:joining}. If experiments are started using \emph{forever} then the server simply restarts the experiment after each session, and the experiment continues until the researcher manually stops it. See \url{https://github.com/foreverjs/forever}
  \item[Bootstrap:] \Cree uses the CSS framework \emph{Bootstrap}, which insures that pages scale seamlessly across small and big screens. In addition it ensures that pages look identical across different web browsers and devices, for given screen sizes. See \url{http://getbootstrap.com}
  \item[jQuery:] This library extends JavaScript and provides additional useful functions to change and manipulate the web pages that subjects see. More specifically it is used to load new pages and collect inputs from subjects. See \url{https://jquery.com}
  \item[D3.js:] \Cree leverages the JavaScript library \emph{D3.js} to build graphics and figures. \emph{D3.js} is used for both the static graphics and the graphical interfaces that subjects interact with and provide inputs through. See \url{http://d3js.org}
\end{description}

Cree is installed on a computer with Node.js using its package manager (\emph{npm}). To install Cree open \emph{Command Prompt} on Windows, or \emph{Terminal} on Mac/Linux, enter \texttt{npm~install~Cree} and press enter. All of the dependencies are automatically installed, or come pre-bundled, with \Cree.
